---
title: "After *SFFA v. Harvard*: Predictions that Didn't Come True"
author: "David Kane"
date: last-modified
published-title: Date
# NOTE: the abstract lives in YAML, so its numbers CANNOT be computed
# inline with R and are hard-coded by necessity. If clean.csv is updated,
# re-verify every figure below against the rendered body (especially the
# claims keyed to the 33 colleges, the research universities, and
# Harvard's 2010-2015 level).
abstract: |
  A landmark 29 June 2023 U.S. Supreme Court decision (*SFFA v.
  Harvard*) held that race-based admissions at colleges and universities
  were unconstitutional. During the case, elite schools made predictions
  about what would happen if SFFA were to win (as it went on to do).
  Thirty-three selective schools predicted Black enrollment declines of
  50 to 70 percent. Harvard's expert witness, David Card, estimated that
  eliminating racial preferences would cut the Black share of Harvard's
  admitted class from 14 percent to 6 percent. Years have passed since
  the 2023 Supreme Court decision. Here, I test the predictions using
  Common Data Set and IPEDS records assembled for sixty-three selective
  institutions through fall 2025. Once the incompatible definitions of
  population and race underlying the litigation figures are reconciled,
  I find that: (1) Average Black enrollment among the 33 colleges
  exceeds its pre-litigation benchmark. (2) Black enrollment at the
  research universities is essentially unchanged. (3) Harvard's Black
  enrollment matches its 2010-2015 level. The results raise the issue:
  Does abiding by the decision not have the predicted effect, or are the
  schools not abiding by the decision?

  *Keywords*: affirmative action, college admissions, expert testimony,
  Students for Fair Admissions, Common Data Set.

  *JEL codes*: I23, I28, J7.
fig-cap-location: top
tbl-cap-location: top
number-sections: false
format:
  html:
    toc: true
    embed-resources: true
  pdf:
    toc: false
  docx:
    toc: false
    reference-doc: ejw-reference.docx
---

```{r}
#| label: setup
#| include: false
library(dplyr)
library(knitr)
library(ggplot2)
library(scales)
cds <- read.csv("../data/processed/clean.csv", check.names = FALSE)

# School groups, defined once here so that counts quoted in early prose
# (e.g. the Data section's filing coverage) can be computed inline.
FIFTEEN <- c("brown","caltech","cmu","columbia","cornell","dartmouth",
             "duke","emory","jhu","princeton","penn","uchicago",
             "vanderbilt","wustl","yale")
AMICI <- c("amherst","barnard","bates","bowdoin","brynmawr","carleton",
           "colby","conncoll","davidson","fandm","hamilton","hampshire",
           "haverford","macalester","middlebury","mtholyoke","oberlin",
           "pomona","reed","sarahlawrence","smith","stolaf","swarthmore",
           "trinity","union","vassar","wellesley","williams",
           "bucknell","clark","tufts","washandlee","wesleyan")
US_NEWS_UNIVS <- c(
  "princeton", "mit", "harvard", "stanford", "yale", "uchicago",
  "duke", "jhu", "northwestern", "penn", "caltech", "cornell",
  "brown", "dartmouth", "columbia", "berkeley", "rice", "ucla",
  "vanderbilt", "cmu", "umich", "notredame", "wustl", "emory",
  "georgetown")
# Service academies (Navy, Air Force, Army) are ranked in the US News
# top-25 LACs but excluded from the analysis.
US_NEWS_LACS <- c(
  "williams", "amherst", "swarthmore", "bowdoin",
  "cmc", "pomona", "wellesley", "carleton", "hmc",
  "barnard", "davidson", "grinnell", "hamilton", "middlebury",
  "smith", "vassar", "wesleyan", "washandlee", "colgate", "richmond",
  "bates", "colby")
TOPRANKED <- c(US_NEWS_UNIVS, US_NEWS_LACS)
SCHOOLS63 <- union(c("harvard", FIFTEEN, AMICI), TOPRANKED)
stopifnot(length(SCHOOLS63) == 63)

# Fall 2025 filing coverage. "Usable" means the filing reports a Black
# first-year count (Bryn Mawr's 2025-26 CDS omits the racial table and
# is therefore not usable; Macalester's reports race but no Total row
# and counts as usable here).
has_2025_black <- cds |>
  filter(year == 2025,
         category == "Black or African American, non-Hispanic") |>
  mutate(n = suppressWarnings(as.numeric(first_time_first_year))) |>
  filter(!is.na(n)) |>
  pull(school)
n2025_63 <- sum(SCHOOLS63 %in% has_2025_black)
n2025_47 <- sum(TOPRANKED %in% has_2025_black)

harvard <- cds |>
  filter(school == "harvard") |>
  mutate(n = suppressWarnings(as.numeric(first_time_first_year))) |>
  filter(!is.na(n)) |>
  select(year, category, n)

wide <- harvard |>
  tidyr::pivot_wider(names_from = category, values_from = n) |>
  rename(black   = `Black or African American, non-Hispanic`,
         total   = `Total`,
         nonres  = `Nonresident`) |>
  mutate(us_total    = total - nonres,
         black_share = black / us_total)
```

The Supreme Court's decision in *Students for Fair Admissions v. Harvard*
(600 U.S. 181, decided June 29, 2023) prohibited the consideration of race
in college admissions. During the litigation, the affected institutions
committed themselves to forecasts as to what such a
prohibition would do. A coalition of 33 selective colleges filed an amicus
brief insisting that Black enrollment would fall by 50 to 70 percent
(Amherst Brief 2022, 13-14). Harvard expert David Card simulated race-blind
admission for the Class of 2019 and reported that the Black
share of admitted students would fall from 14 percent to 6 percent (Card
2017, 103).

Enrollment records for the first two
admissions cycles conducted under the new legal regime, the entering
classes of fall 2024 and fall 2025, are now available. These forecasts can
therefore be evaluated against realized outcomes, a test that, to my
knowledge, no published study has yet performed.

The evaluation is less straightforward than it appears. The only publicly available enrollment statistics
come from the Common Data Set (CDS), a standardized annual
disclosure filed by nearly every selective college, and from the federal
IPEDS Fall Enrollment survey. Card's figures were computed on a different
population (admitted rather than enrolled students, domestic applicants
only) under a different racial classification scheme (Harvard's internal
coding, which assigns multiracial applicants to single categories). The
same student body therefore carries two Black shares that differ by
roughly a factor of two depending on which convention is applied. Once the expert-report metric and the federal metric
are placed on a common basis, Harvard's post-ruling classes are
indistinguishable from the classes it enrolled while
practicing the admissions system it had prior to the decision.

The paper's primary conclusion is that elite institutions enroll
classes whose racial composition is essentially
unchanged by *SFFA*, a stability their own litigation forecasts said
race-neutral admissions could not produce. I support this conclusion
with several bodies of evidence.

<!-- The numbers in the four evidence paragraphs below are hard-coded
     because knitr computes the underlying objects (rtab, tbl, pool50)
     in chunks that run later in the document. Keep them in sync with
     the Research Universities, Amici, and Top-Ranked Cohort sections,
     and revisit if clean.csv is updated. -->
- Mean Black enrollment across the 33 amici colleges is above,
  not below, its pre-litigation benchmark.

- Mean Black enrollment at the 15 research universities that told the
  Court race-neutral admissions would undercut their diversity moved from
  7.4 to 7.2 percent of US first-years, and six of the fifteen enroll
  more Black students, relative to class size, than before.

- The ratio of Asian to Black first-year students across
  the 47 top-ranked institutions has moved a small fraction of the
  distance toward the levels observed at universities bound by California's
  Proposition 209, the closest available approximation to a race-blind benchmark.

- Black enrollment at Harvard, correctly converted, sits at its 2010-2015
  level.

These results raise the issue: Does abiding by the decision not have
the predicted effect, or are the schools not abiding by the decision?

The paper also raises concerns about the role of economic expertise. Card's
report was produced for litigation and stated its predictions in units
that correspond to no publicly reported statistic. The analyses on the
plaintiff's side did enter scholarly journals (Arcidiacono, Kinsler, and
Ransom 2022a, 2022b), but the defense simulation by Card that became the
public reference point did not, and no mapping between its conventions
and the CDS categories has previously been published. Sixteen prominent
economists—George A. Akerlof, Sandy Baum, Rebecca M. Blank, Hilary
Hoynes, Guido W. Imbens, Kirabo Jackson, Rucker Johnson, Helen F. Ladd,
David S. Lee, Trevon D. Logan, Michael McPherson, Jesse Rothstein,
Robert M. Solow, Sarah Turner, Ebonya Washington, and Douglas
Webber—three of them Nobel laureates, nonetheless attested to the Court
that the analysis was methodologically sound (Economics Professors Brief
2022). The most widely cited prediction in the case is one that the
public record cannot check. Institutions with an interest in appearing
compliant benefit from the resulting ambiguity. Harvard's own use of
this ambiguity in its post-ruling disclosures is documented below. The
conversion that I develop here may help in mitigating the ambiguity.

## Related Literature

A substantial empirical literature, extending back three decades,
estimates the magnitude of racial preferences at selective institutions,
and its results are consistent enough to summarize briefly. William G.
Bowen and Derek Bok, in work first published in 1998, calculated from the 1989 entering
cohort of the College and Beyond schools that race-neutral admission
would have reduced the Black share of matriculants from 7.1 percent to
between 2.1 and 3.6 percent (Bowen and Bok 2018, 39). Espenshade and Chung (2005), simulating admissions at
three elite private universities from 1997 application records, found
that eliminating racial preferences would cut the acceptance rate of
Black candidates from 33.7 to 12.2 percent, reduce the
Black share of the admitted class from 9.0 to 3.3 percent,
and leave Asian applicants filling nearly four of every five vacated
seats. Arcidiacono and Lovenheim (2016) survey the broader literature
and report similar effect sizes wherever selective admissions
are modeled. These simulations are the direct intellectual ancestors of
the forecasts the amici colleges placed before the Supreme Court.

The SFFA litigation itself generated the most detailed estimates in the
literature, because discovery gave the experts on both sides access to
complete applicant-level records. From the Harvard and UNC data,
Peter Arcidiacono, Josh Kinsler, and Tyler Ransom published three
peer-reviewed studies.
They found that the typical Black admit to Harvard would have
faced an admission rate roughly four times lower had the identical
application been evaluated as White, with preferences largest for
applicants from advantaged backgrounds (Arcidiacono, Kinsler, and Ransom
2023), that Asian American applicants faced a penalty relative to
observably similar White applicants (Arcidiacono, Kinsler, and Ransom
2022a), and that preferences for legacies and athletes flowed
overwhelmingly to White applicants (Arcidiacono, Kinsler, and Ransom
2022b). Card's simulated decline from 14 to 6 percent, described in the
next section, is the defense counterpart of these estimates. The two
sides disputed modeling choices at length, but no expert in the case
produced a model in which removing racial preferences left the racial
composition of the class approximately unchanged. 

A complementary quasi-experimental literature studies what happens when
selective institutions actually stop considering race. California's
Proposition 209 is the cleanest case. Bleemer (2022) links every
University of California applicant from 1994 through 2002 to later
educational and labor-market records and shows that the ban produced
immediate, large declines in underrepresented minority enrollment at the
most selective campuses, cascading those applicants into less selective
institutions and lowering their subsequent wages. Hinrichs (2012) finds
in national data that statewide bans leave aggregate minority enrollment
roughly unchanged while reducing it sharply at the most selective
institutions, the segment occupied by every school studied here. These
episodes supply an empirical benchmark. Where
race-conscious admission genuinely ends at a highly selective
institution, the compositional change is large and immediate.

The stakes of the compositional question are documented by Chetty,
Deming, and Friedman (2026), who estimate the causal effect of
attending an Ivy-Plus college on reaching the top of the income
distribution and on entry into leadership positions. Their subject is
the consequence of admission rather than its determinants, but their
findings explain why the demographics of a few dozen institutions
enrolling a fraction of one percent of American undergraduates command
the attention they do. The first post-ruling cycle has begun to attract
systematic study. Bloem et al. (2026) link national testing cohorts to
fall 2024 enrollment records and document underrepresented minority
enrollees cascading down the selectivity distribution in the ruling's
first year. Their estimates are measured against
the fall 2021-2023 entering classes, the elevated post-2020 cohorts
examined below. Measured against the 2010-2015 litigation-era baseline,
the same institutions show the stability documented in this paper. What no part of this literature yet
contains is a comparison of the *SFFA* predictions with realized
post-ruling enrollment, stated in the units in which the predictions were issued and measured against the baseline the predictions themselves referenced.
The prediction side of the record closed in 2023. The outcome side, now
two entering classes deep, sits in public filings, and the remainder of
this paper performs the comparison.

## The Litigation Forecasts

Card's report, filed December 15, 2017, estimated admissions models on
Harvard's internal records for applicants to the Classes of 2014 through
2019 and then simulated the Class of 2019 admissions cycle with the race
coefficients set to zero:

> In my simulation, the share of African-American students in the admitted
> class would drop from 14% to 6%. The fraction of Hispanic or Other students
> would fall from 14% to 9%. The fraction of admitted students who are
> Asian-American would rise from 24% to 27%. And the fraction of admitted
> students who are White would rise from 40% to 48%.
> (Card 2017, 103)

The simulation's baseline of 14 percent has a specific and unfamiliar
meaning. Card's sample contains the 1,679 admitted students to the Class
of 2019 who appear in his preferred regression model, and 234 of them,
or 14 percent, are recorded as African American (Card 2017, 125). Three
restrictions separate this population from anything in a public filing.
It counts admitted students rather than enrolled ones. It is limited to
domestic applicants, defined as citizens and permanent residents, a
restriction that runs throughout the report (Card 2017, nn. 6, 8). And
it classifies race using Harvard's internal scheme, which assigns an
applicant marking more than one race to a single category, so Card's
Black count is broader than the single-race Black line in a federal
filing. Harvard admitted roughly 2,000 students in the cycle, and the
several hundred missing from the sample are internationals and
applicants outside the regression model. None of this makes the sample
unusable as a baseline. The conversion developed below does not require
Card's sample to match the enrolled class member for member, only that
the relationship between his population and the enrolled class be
stable, an assumption that cannot be verified from public data and
whose plausible failures, examined below, mostly run against the paper's
conclusion.

Two amicus briefs, both filed August 1, 2022, carried versions of these
forecasts to the Supreme Court, and their quantitative commitments differ
markedly. The first came from fifteen research universities
(Brown, Caltech, Carnegie Mellon, Columbia, Cornell, Dartmouth, Duke,
Emory, Johns Hopkins, Princeton, Penn, the University of Chicago,
Vanderbilt, Washington University in St. Louis, and Yale), a group that
includes neither Harvard nor any signatory of the second brief. Its
one explicit number is Harvard-specific: the trial record showed that the
race-neutral alternatives proposed by the petitioner "would lead to a near
33% reduction in the number of African American students admitted" at
Harvard (Universities Brief 2022, 25, quoting the district court's
findings). The finding the brief quotes summarizes the race-neutral
simulations offered by the petitioner's expert, under which African
American representation in Harvard's incoming class "would fall nearly
one-third to approximately 10% of the class" (*SFFA v. Harvard*, 397
F. Supp. 3d at 178-84). The 33 percent figure is smaller than Card's 57 percent
relative decline because it describes a different counterfactual, the
petitioner's proposed alternatives with their compensating policy changes,
rather than the removal of race with nothing else altered. Both figures
rest on the same underlying applicant data. Regarding their own campuses, the
fifteen universities made no numerical forecast. Their claims were
qualitative but unambiguous in direction: exclusively race-neutral
admissions "would undercut Amici's efforts to attain 'the benefits of
diversity' they seek," and "no race-neutral alternative presently can
fully replace race-conscious individualized and holistic review"
(Universities Brief 2022, 25, 31). The brief thus committed its
signatories to the proposition that compliance with the petitioner's
requested rule would visibly reduce racial diversity at their own
institutions.

The second brief, organized by Amherst and joined by thirty-two other
colleges and universities, made the broadest quantitative claims in the record:

> Black enrollment would likely decline between 50% and 70%; the
> probability of Black applicants receiving offers of admission would drop
> to half that of white students; and the percentage of Black students
> matriculating would drop from roughly 7.1% of the student body to 2.1%.
> Strictly enforced, a 'do not consider race' policy would 'presumably take
> [B]lack enrollments . . . back to early 1960s levels, before colleges and
> universities began to make serious efforts to recruit minority
> students.'
> (Amherst Brief 2022, 13-14)

The brief's footnote for the entire quoted passage is Bowen and Bok (2018,
39), whose simulations on the 1989 entering cohort of the College and
Beyond institutions had Black matriculants falling from 7.1 percent to
between 2.1 and 3.6 percent, declines of 49 to 70 percent. The interval,
the halved admission probability, and the 7.1-to-2.1 endpoints are
therefore one simulation, run on a cohort that entered college a third of
a century before the ruling. A separate footnote attaches Grodsky and
Kurlaender (2006, 7-10) to the preceding sentence's general claim of a
"drastic resegregating impact." The single
contemporary, institution-specific estimate in the brief is attributed to
Amherst's internal planning: an entirely race-blind policy would reduce
the college's combined Native American, Black, and Hispanic enrollment "by
approximately half" (Amherst Brief 2022, 14). The supporting footnote
supplies neither a document, a date, nor a cohort, noting only that the
underlying studies formed part of the periodic review of admissions policy
that *Grutter v. Bollinger* requires of institutions using racial
preferences. The amici therefore left no single well-specified
counterfactual, and I evaluate their forecast as a proportional claim
against several candidate baselines below.

The economics profession participated in the litigation directly, on both
sides. A group of economists filed in support of the petitioner attacking
the lower courts' reliance on Card's model (Economists Brief 2022).
Sixteen economists and statisticians answered in support of Harvard,
among them three Nobel laureates, George Akerlof, Robert Solow, and Guido
Imbens, the last a co-recipient of the 2021 prize that Card
shared. Their brief attested that Card "is one of the most outstanding
and respected scholars in the field of econometrics and applied
economics," that "his statistical analyses in this case were
methodologically sound," and that "the criticisms of the lower courts'
consideration of Dr. Card's analyses in the Brief of Economists as Amici
Curiae in Support of Petitioner . . . are not credible" (Economics
Professors Brief 2022, 1-2).
Prominent economists endorsed Card's analysis to the Court, so the predictions tested below
command professional endorsement and are not simply one paid expert's judgment.

## Post-SFFA Results

### Data

I assembled enrollment records for sixty-three selective colleges and
universities from two sources. The sixty-three are the union of the
forty-seven institutions in the 2026 US News top-25 National University
and top-25 National Liberal Arts College lists with the forty-nine
litigation participants examined above (Harvard, the fifteen
universities, and the thirty-three amici); the two groups share
thirty-three members. The three federal service academies appear in the
liberal arts ranking but are excluded throughout: their admissions run
through congressional nomination and service commitments, and the Naval
Academy publishes no Common Data Set at all. The first
source is CDS item B2, which reports
first-time first-year enrollment by racial and ethnic category. I
collected every publicly available CDS filing for each institution from
1998-99 through 2025-26, extracting the B2 tables from PDF, spreadsheet,
and web formats. The second is the IPEDS Fall Enrollment component for
2008 through 2024, which reports the same categories under the same
federal definitions. For any institution-year covered by both sources, the
IPEDS record is used, since its uniform structure eliminates the parsing
errors that school-specific PDF layouts occasionally introduce. CDS
filings supply the years outside IPEDS coverage, most importantly fall
2025, for which `r xfun::n2w(n2025_63)` of the sixty-three institutions had published a
usable filing at the time of writing (`r xfun::n2w(n2025_47)` of them among the
forty-seven top-ranked institutions, the count relevant to the cohort figures
below). All statistics below refer to first-time first-year students.

Two definitions recur. A "US" share divides by total first-year enrollment
net of nonresident students, who appear on a separate B2 line and receive
no racial classification in these data. A "whole-class" share divides by
total first-year enrollment. Federal race categories follow the 2010 IPEDS
rules. Hispanic students of any race are reported as Hispanic,
non-Hispanic students reporting two or more races are reported in a
dedicated multiracial category, and the Black line therefore contains only
single-race, non-Hispanic Black students. These rules differ substantially
from the classification underlying Card's figures, a difference quantified
below.

### Harvard

#### Commensurability

The three restrictions described in the forecasts section mean that Card's 14 percent
corresponds to no number in a public filing. The federal side of the
comparison has its own conventions. A CDS filing counts enrolled
students, roughly 1,600 in Harvard's case, against the roughly 2,000
admitted. It reports international students on a separate line of the
same table rather than excluding them. And it applies the federal race
rules, under which Hispanic students of any race are reported as
Hispanic and non-Hispanic multiracial students occupy a dedicated
category, so the single-race Black line is systematically narrower than
Card's Black count.
Matriculation rates do vary somewhat by race, with Black admits enrolling
at moderately lower rates than White or Asian admits according to the
trial record (*SFFA v. Harvard*, 397 F. Supp. 3d at 155-56), but the
magnitude is small relative to the classification wedge, and I treat the
combined gap as a single conversion to be estimated rather than
decomposed.

The estimation exploits the fact that both metrics can be observed for the
identical cohort, the one cohort for which this is possible. Card's
baseline describes students admitted in the
spring of 2015. Those who enrolled constitute the fall 2015 entering
class, which appears in Harvard's CDS 2015-16 filing as 109 single-race
Black first-years among 1,441 US first-years, or 7.56 percent. The ratio
of the federal figure to the expert-report figure, computed on the
underlying counts (109/1,441 divided by 234/1,679), is 0.543, and I use
0.54 below. Throughout
the paper, shares are displayed to one decimal and proportional changes
to the nearest percent, and all calculations use unrounded values.

```{r}
#| include: false
# The conversion factor is calibrated on the single cohort observable in
# both metrics: the Class of 2019, admitted spring 2015, entering fall
# 2015. Card's side is 234/1,679 (Card 2017, 125); the CDS side is
# Harvard's fall 2015 filing, taken from clean.csv. base_share (the fall
# 2010-2015 CDS mean) anchors the CDS-side pre/post comparison only.
baseline <- wide |> filter(year %in% 2010:2015)
base_share <- mean(baseline$black_share)
card_cal <- 234 / 1679
cds_cal <- wide$black_share[wide$year == 2015]
factor <- cds_cal / card_cal    # CDS units per Card unit
```

A prediction stated in Card's units is
converted to the federal metric by multiplication by this factor, so the
simulated race-blind share of 6 percent corresponds to
`r sprintf("%.1f", 100 * 0.06 * factor)` percent on a CDS filing. This
converted benchmark, rather than the unconverted 6 percent, is the
quantity against which post-ruling filings should be compared.

<!-- Evidence for "the yield differences documented at trial amounted to
     a few percentage points" (below): the district court's findings,
     397 F. Supp. 3d at 155-56 (ECF Doc. 672 at 29), state the DIRECTION
     only: admitted Asian American students matriculate at a higher rate
     than White students, and Hispanic, African American, Native
     American, and multiracial admits at lower rates, citing trial
     exhibit PX324 and Fitzsimmons testimony (Oct. 18 Tr. 80:21-81:5).
     PX324 holds the actual rates and is not in our documents/
     collection. "A few percentage points" is our bounding
     characterization of the channel's effect on shares: Harvard's
     overall yield is roughly 80 percent (about 2,000 admits for a class
     of about 1,600, 397 F. Supp. 3d at 134), and even a 10-point lower
     Black yield would move the Black share of enrollees only ~1.5
     percentage points below the admit share (0.139*0.70 /
     (0.139*0.70 + 0.861*0.80) = 12.4% vs 13.9%), second order against
     the ~2x classification wedge. -->
The conversion factor bundles the multiracial and Hispanic
reassignments, the admit-to-enrollee step, and the domestic-scope
adjustment, all measured on the fall 2015 cohort, and may have drifted
since. National rates of multiracial identification have risen
since then, which would push the current factor below 0.54 and the
converted benchmark below 3.3 percent. Nondisclosure of race has also
risen, which depresses measured single-race shares in recent filings.
Both developments bias the comparison toward finding compliance.
Matriculation rates are a third potential source of drift, since
post-ruling recruiting or financial aid could have changed the rate at
which admitted Black students enroll, but the yield differences
documented at trial amounted to a few percentage points against a
classification wedge of nearly a factor of two, so movement in this
channel is second order. Estimates
of no change therefore cannot be attributed to drift in the factor, a
point developed further below.

#### The Post-Ruling Classes

Harvard's CDS filing for 2025-26 reports 97 Black first-year students in
an entering class of 1,673, which is 5.8 percent of the whole class and
6.8 percent of US first-years. Read against the unconverted simulation, a
figure of 5.8 percent suggests that Card's race-blind counterfactual has
essentially been realized. Read against the converted benchmark of 3.3
percent, it indicates the opposite. @tbl-harvard assembles the comparison in
both metrics, averaging the two post-ruling classes to reduce
year-to-year noise, to parallel the 2010-2015 averaging of the
baseline.

```{r}
#| echo: false
#| label: tbl-harvard
#| tbl-cap: "Black share of Harvard's entering class under the federal (CDS) metric and the expert-report (Card) metric. The CDS metric is the single-race Black share of enrolled US first-years. The conversion factor between metrics is 0.54, calibrated on the Class of 2019, the one cohort observable in both metrics. Card-unit entries other than the calibration cohort are obtained by conversion; the fall 2010-2015 average is a CDS-side quantity only."
post <- wide |> filter(year %in% 2024:2025)
post_share <- mean(post$black_share)
tibble::tibble(
  ` ` = c("Calibration cohort (Class of 2019, entered fall 2015)",
          "Fall 2010-2015 average",
          "Race-blind simulation (Card 2017)",
          "Post-ruling classes (Classes of 2028-29, avg)"),
  `CDS units`  = sprintf("%.1f%%", 100 * c(cds_cal, base_share, 0.06 * factor, post_share)),
  `Card units` = c(sprintf("%.1f%%", 100 * card_cal), "—",
                   sprintf("%.1f%%", 100 * c(0.06, post_share / factor)))
) |> kable(align = "lrr")
```

The entering classes of fall 2024 and fall 2025 carried US Black shares of
8.14 and 6.82 percent respectively, averaging
`r sprintf("%.2f", 100*post_share)` percent. The 2010-2015 mean is
`r sprintf("%.2f", 100*base_share)` percent. The post-ruling average sits
within about a twentieth of a percentage point of the pre-ruling average, and
both display as 7.5 percent at the paper's reporting precision. Converted
into Card's own units, the post-ruling classes correspond to
`r sprintf("%.1f", 100*post_share/factor)` percent Black, against the
calibration cohort's 13.9. If the
race-blind counterfactual in the expert report is taken at face value, the
gap between what a compliant Harvard would report and what Harvard does
report exceeds a factor of two.

Two measurement objections require examination. The first concerns
nondisclosure. If Black students became disproportionately likely to
decline the race question after the ruling, single-race Black shares would
fall mechanically. The denominator of the US share includes
nondisclosers, so a withheld identification subtracts from the numerator
while leaving the denominator intact, and the bias can only run
downward. The direction of this bias runs toward compliance, not
away from it, so it cannot generate a spurious finding of no change. Its
scale is also limited. Harvard's unknown-race share averaged 5.3 percent
over the 2010-2015 baseline cohorts, reached a low of 2.4 percent in
2023, and averages near 8 percent in the post-ruling classes. Relative to
the calibration period, the increase is under three percentage points,
and Card's own sample carried an 8 percent race-missing rate. The second
objection concerns drift in the conversion factor, addressed above:
plausible drift lowers the converted benchmark and widens, rather than
narrows, the measured gap. In summary, the filings contain no evidence
of movement toward race-blind admissions.

#### Post-Ruling Disclosure Practice

The unit ambiguity documented above is not confined to expert
reports. Harvard's own public communications since the ruling have
employed a metric distinct from the one in its federal filings. The
announcement for the fall 2024 class stated that 14 percent of students
identifying a race identified as African American or Black (Harvard
Gazette 2024). Two definitional choices produced that figure: race is
self-identified without single-race restriction, and the denominator
excludes the nondisclosing group, which doubled to 8 percent that same
year. The corresponding IPEDS figure for the identical class is 8.1
percent. For the fall 2025 class, the university moved its demographic
announcement to late October, when federal reporting obligations come due,
and reported 11.5 percent on the same disclosers-only basis (Harvard Crimson
2025), against 6.8 percent of US first-years in its CDS filing. The two
sources describe the same students, 1,675 enrolling students in the
announcement against 1,673 first-time first-years in the filing, a
difference presumably attributable to the timing of the fall census. An
institution whose public number reproduces its litigation-era level while
its federal number halves that level enjoys considerable interpretive
latitude. Where the categories of expert analysis, public
communication, and
statutory reporting fail to correspond, enrollment outcomes cannot be
monitored by inspection, and claims of compliance or of catastrophe can
each find a supporting statistic.

### The Research Universities

The fifteen universities made no numerical forecast about themselves, but
their qualitative commitment is testable. If
exclusively race-neutral admissions would undercut the racial diversity
these institutions seek, then compliance with the ruling should have
produced visible declines in Black enrollment on their own campuses.
@tbl-fifteen reports Black shares of US first-years, the same metric used for
Harvard, over the same pre-period, for each signatory.

```{r}
#| echo: false
#| label: tbl-fifteen
#| tbl-cap: "Black share of first-year US students at the fifteen research universities that filed the Universities Brief, ordered by proportional change. The pre-period comprises the entering classes of fall 2010-2015. The post-period averages fall 2024 and fall 2025 where both filings exist. Changes are computed on unrounded shares and may differ slightly from values recomputed from the rounded shares shown."
# FIFTEEN is defined in the setup chunk.
UNAMES <- c(brown = "Brown", caltech = "Caltech",
  cmu = "Carnegie Mellon", columbia = "Columbia", cornell = "Cornell",
  dartmouth = "Dartmouth", duke = "Duke", emory = "Emory",
  jhu = "Johns Hopkins", princeton = "Princeton", penn = "Penn",
  uchicago = "Chicago", vanderbilt = "Vanderbilt",
  wustl = "Washington Univ.", yale = "Yale")

ru <- cds |>
  filter(school %in% FIFTEEN,
         category %in% c("Black or African American, non-Hispanic",
                         "Total", "Nonresident")) |>
  mutate(n = suppressWarnings(as.numeric(first_time_first_year))) |>
  select(school, year, category, n) |>
  tidyr::pivot_wider(names_from = category, values_from = n) |>
  rename(black  = `Black or African American, non-Hispanic`,
         total  = `Total`,
         nonres = `Nonresident`) |>
  mutate(nonres = ifelse(is.na(nonres), 0, nonres),
         share  = black / (total - nonres))

rp <- function(yrs, name) {
  ru |> filter(year %in% yrs) |> group_by(school) |>
    summarise("{name}" := mean(share, na.rm = TRUE), .groups = "drop")
}

rtab <- rp(2010:2015, "litigation") |>
  left_join(rp(2024:2025, "post"), by = "school") |>
  mutate(change = post / litigation - 1) |>
  arrange(change)

rt_pct <- function(x) ifelse(is.nan(x) | is.na(x), "—",
                             sprintf("%.1f%%", 100*x))
rt_chg <- function(x) {
  p <- round(100 * x)
  ifelse(is.nan(x) | is.na(x), "—",
         ifelse(p == 0, "0%", sprintf("%+.0f%%", p)))
}

r_avg <- rtab |>
  summarise(school = "all",
            litigation = mean(litigation, na.rm = TRUE),
            post = mean(post, na.rm = TRUE)) |>
  mutate(change = post / litigation - 1)

bind_rows(rtab, r_avg) |>
  transmute(University = ifelse(school == "all",
                                "All 15 signatories (mean)",
                                UNAMES[school]),
            `Fall 2010-15 avg` = rt_pct(litigation),
            `Fall 2024-25 avg` = rt_pct(post),
            `Change` = rt_chg(change)) |>
  kable(align = "lrrr")
```

<!-- "decline" below assumes r_avg$change is negative; the school names
     (Carnegie Mellon, Caltech, Yale, WashU, Penn) and the claims "no
     signatory sits at half its baseline" / "a single one" are hand-checked
     against tbl-fifteen. Revisit if clean.csv is updated. -->
The group mean moved from
`r sprintf("%.1f", 100*mean(rtab$litigation, na.rm=TRUE))` percent to
`r sprintf("%.1f", 100*mean(rtab$post, na.rm=TRUE))` percent, a relative
decline of about `r sprintf("%.0f", abs(100*r_avg$change))` percent.
`r xfun::n2w(sum(rtab$change > 0, na.rm = TRUE), cap = TRUE)` of the fifteen enroll a higher Black
share than in the pre-period. No signatory sits at half its baseline, and
a single one, Carnegie Mellon, shows a decline exceeding even the 33
percent figure the brief quoted for a milder counterfactual at Harvard.
Caltech's large proportional increase reflects a very low base, from
`r sprintf("%.1f", 100*rtab$litigation[rtab$school == "caltech"])`
to `r sprintf("%.1f", 100*rtab$post[rtab$school == "caltech"])` percent,
and Yale, Washington University, and Penn all enroll
between `r {three <- rtab$change[rtab$school %in% c("yale", "wustl", "penn")]; sprintf("%.0f and %.0f", 100*min(three), 100*max(three))}` percent more Black first-years, relative to class size,
than they did in the pre-period. If the signatories complied with the
ruling, their own filings refute their claim that compliance would
visibly reduce their racial diversity. 

### The Amici Colleges

The amici forecast refers to "the student body" without restriction to
domestic students and without a stated racial classification, so I
evaluate it in whole-class CDS shares, which require no conversion. Doing
so is conservative. International enrollment at these
colleges grew from roughly 9 percent of the entering class in the early
2010s to roughly 11 percent after the ruling, mechanically deflating
recent whole-class shares relative to older ones. Measured stability
therefore understates underlying stability.

```{r}
#| include: false
# AMICI is defined in the setup chunk.
ab <- cds |>
  filter(school %in% AMICI,
         category %in% c("Black or African American, non-Hispanic",
                         "Total")) |>
  mutate(n = suppressWarnings(as.numeric(first_time_first_year))) |>
  select(school, year, category, n) |>
  tidyr::pivot_wider(names_from = category, values_from = n) |>
  rename(black = `Black or African American, non-Hispanic`,
         total = `Total`) |>
  mutate(share = black / total)   # whole class, as the brief measures it

period_avg <- function(yrs, name) {
  ab |> filter(year %in% yrs) |> group_by(school) |>
    summarise("{name}" := mean(share, na.rm = TRUE), .groups = "drop")
}

tbl <- period_avg(2010:2015, "litigation") |>
  left_join(period_avg(2020:2023, "floyd"),  by = "school") |>
  left_join(period_avg(2024,      "y2024"),  by = "school") |>
  left_join(period_avg(2025,      "y2025"),  by = "school") |>
  left_join(period_avg(2024:2025, "post"),   by = "school") |>
  mutate(change = post / litigation - 1) |>
  arrange(change)

# Helpers for inline prose below.
half_n <- sum(tbl$post <= 0.5 * tbl$litigation, na.rm = TRUE)
base_ratio <- function(s) tbl$post[tbl$school == s] / tbl$litigation[tbl$school == s]
hv2025 <- ab |> filter(school == "haverford", year == 2025)
```

@tbl-amici reports whole-class Black shares for every signatory, ordered by
proportional change, together with the group mean. The pre-period matches
the Harvard estimation window, the entering classes of fall 2010 through
fall 2015. Post-ruling entries reflect fall 2024 alone where no usable
2025-26 filing exists. 

<!-- Bryn Mawr published a 2025-26 CDS that omits the racial composition table.-->


```{r}
#| echo: false
#| label: tbl-amici
#| tbl-cap: "Whole-class Black share for all thirty-three signatories of the Amherst brief, ordered by proportional change. The pre-period comprises the entering classes of fall 2010-2015. The post-period averages fall 2024 and fall 2025, or reflects fall 2024 alone where no usable 2025-26 filing exists. Changes are computed on unrounded shares and may differ slightly from values recomputed from the rounded shares shown."
NAMES <- c(amherst = "Amherst", barnard = "Barnard", bates = "Bates",
  bowdoin = "Bowdoin", brynmawr = "Bryn Mawr", carleton = "Carleton",
  colby = "Colby", conncoll = "Connecticut College", davidson = "Davidson",
  fandm = "Franklin & Marshall", hamilton = "Hamilton",
  hampshire = "Hampshire", haverford = "Haverford",
  macalester = "Macalester", middlebury = "Middlebury",
  mtholyoke = "Mount Holyoke", oberlin = "Oberlin", pomona = "Pomona",
  reed = "Reed", sarahlawrence = "Sarah Lawrence", smith = "Smith",
  stolaf = "St. Olaf", swarthmore = "Swarthmore", trinity = "Trinity",
  union = "Union", vassar = "Vassar", wellesley = "Wellesley",
  williams = "Williams", bucknell = "Bucknell", clark = "Clark",
  tufts = "Tufts", washandlee = "Washington & Lee", wesleyan = "Wesleyan")

pct <- function(x) ifelse(is.nan(x) | is.na(x), "—", sprintf("%.1f%%", 100*x))
chg <- function(x) {
  p <- round(100 * x)
  ifelse(is.nan(x) | is.na(x), "—",
         ifelse(p == 0, "0%", sprintf("%+.0f%%", p)))
}

fmt <- function(df, label = NULL) {
  transmute(df,
    School = if (is.null(label)) NAMES[school] else label,
    `Fall 2010-15 avg` = pct(litigation),
    `Fall 2024-25 avg` = pct(post),
    `Change` = chg(change))
}

avg_row <- tbl |>
  summarise(school = "all", litigation = mean(litigation, na.rm = TRUE),
            floyd = mean(floyd, na.rm = TRUE),
            y2024 = mean(y2024, na.rm = TRUE),
            y2025 = mean(y2025, na.rm = TRUE),
            post  = mean(post,  na.rm = TRUE)) |>
  mutate(change = post / litigation - 1)

bind_rows(
  tbl |> arrange(change) |> fmt(),
  avg_row |> fmt(label = "All 33 signatories (mean)")
) |> kable(align = "lrrr")
```

Mean Black enrollment across the signatories was
`r sprintf("%.1f", 100*mean(tbl$litigation, na.rm=TRUE))` percent of the
entering class during the pre-period, climbed to
`r sprintf("%.1f", 100*mean(tbl$floyd, na.rm=TRUE))` percent in the
entering classes of 2020 through 2023, and stands at
`r sprintf("%.1f", 100*mean(tbl$post, na.rm=TRUE))` percent in the two
post-ruling classes. Relative to the benchmark period, the change is
positive. The group mean did not fall at all, let alone by half, so the
forecast fails at the level of the group before any school-level
accounting. Because some
institutions lack a usable 2025-26 filing, the post-period mixes one- and
two-year averages. Of the thirty-three,
`r xfun::n2w(sum(!is.na(tbl$y2024) & !is.na(tbl$y2025)))` have filings for both
post-ruling years, and the mean over that balanced panel is
`r sprintf("%.1f%%", 100*mean(tbl$post[!is.na(tbl$y2024) & !is.na(tbl$y2025)]))`,
similar to the full-group figure.

<!-- The school names (Amherst, Haverford, Vassar, Bates), the ordering
     claims, and the windows sentence at the end of this paragraph
     ("two under three of the windows and one under the fourth", computed
     in the hidden windows chunk below, which runs after this prose) are
     hand-checked against the tables. Revisit if clean.csv is updated. -->
Interpreting the brief's 2.1 percent endpoint as an absolute level would
misapply it, since that figure describes Bowen and Bok's 1989 cohort
rather than these colleges, whose own pre-period mean was
`r sprintf("%.1f", 100*mean(tbl$litigation, na.rm=TRUE))` percent.
The transferable content of the forecast is proportional. Applied to each
school's own baseline, even the mildest version implies a post-ruling
share no greater than half the pre-ruling share.
`r xfun::n2w(half_n, cap = TRUE)` signatories out of
thirty-three satisfy this criterion. Amherst's post-ruling average is
`r sprintf("%.2f", base_ratio("amherst"))` of its baseline, and Haverford's is
`r sprintf("%.2f", base_ratio("haverford"))`, a position produced by its
fall 2025 filing, which reports `r xfun::n2w(hv2025$black)` Black first-year students in a
class of `r hv2025$total`. The next closest are Vassar at
`r sprintf("%.2f", base_ratio("vassar"))` and Bates at
`r sprintf("%.2f", base_ratio("bates"))`, and
no other signatory falls below two thirds of its baseline. Because the
choice of pre-period is not dictated by the brief, I repeated the
calculation for four alternative windows, including the complete set of
filings that were public on the day the brief was signed. The group mean
and the implied forecast interval are insensitive to the choice, and the
count of signatories below half their baseline is two under three of the
windows and one under the fourth.

```{r}
#| include: false
# Sensitivity of the amici evaluation to the choice of pre-period,
# summarized in the prose above (table not shown). inv columns: baseline
# mean, post-ruling mean, and the count of signatories at or below one
# half of their own baseline.
windows <- list(`Fall 2010-15 (litigation estimation window)` = 2010:2015,
                `Fall 2016-19`  = 2016:2019,
                `Fall 2020-23 (final preference-era classes)`  = 2020:2023,
                `Fall 2010-21 (all filings public at signing)` = 2010:2021)
inv <- t(sapply(windows, function(yrs) {
  base <- period_avg(yrs, "base"); p <- period_avg(2024:2025, "post")
  x <- inner_join(base, p, by = "school")
  c(base = mean(x$base, na.rm = TRUE), post = mean(x$post, na.rm = TRUE),
    n = sum(x$post <= 0.5 * x$base, na.rm = TRUE))
}))
```

## The Top-Ranked Cohort

Extending the analysis to the union of the 2026 US News top-25 National
Universities and top-25 National Liberal Arts Colleges, less the three
service academies, brings the total to forty-seven institutions. Unlike
the three groups examined so far, these forty-seven never acted
collectively in the litigation, and several filed nothing at
all. The question here is not whether a specific forecast held but whether
the patterns documented for the litigation participants extend across the market segment
they occupy. @fig-black plots the US Black share of each school's
entering class from fall 2006 through fall 2025 together with the
cross-sectional median.

```{r}
#| include: false
# TOPRANKED (US News top-25 universities + top-25 LACs, less the service
# academies) is defined in the setup chunk.
pool50 <- cds |>
  filter(school %in% TOPRANKED, year %in% 2006:2025,
         category %in% c("Black or African American, non-Hispanic",
                         "Asian, non-Hispanic", "Total", "Nonresident")) |>
  mutate(n = suppressWarnings(as.numeric(first_time_first_year))) |>
  select(school, year, category, n) |>
  tidyr::pivot_wider(names_from = category, values_from = n) |>
  rename(black  = `Black or African American, non-Hispanic`,
         asian  = `Asian, non-Hispanic`,
         total  = `Total`,
         nonres = `Nonresident`) |>
  mutate(nonres   = ifelse(is.na(nonres), 0, nonres),
         us_total = total - nonres) |>
  filter(is.finite(us_total), us_total >= 100)

# Tufte-styled spaghetti plot: no grid, recessive axes, direct labels in
# place of a legend, and a dashed rule (rather than a shaded band) marking
# the ruling.
spaghetti50 <- function(df, title, y_format) {
  med <- df |> group_by(year) |>
    summarise(y = median(y, na.rm = TRUE), .groups = "drop")
  med_end <- med |> filter(year == max(year))
  ggplot(df, aes(year, y, group = school)) +
    geom_vline(xintercept = 2023.5, color = "grey60",
               linewidth = 0.4, linetype = "22") +
    geom_line(color = "grey80", linewidth = 0.3, na.rm = TRUE) +
    geom_line(data = med, aes(year, y), inherit.aes = FALSE,
              color = "darkorange", linewidth = 1.2) +
    annotate("text", x = med_end$year + 0.35, y = med_end$y,
             label = "Median", color = "darkorange", fontface = "bold",
             hjust = 0, size = 4.2) +
    annotate("text", x = 2023.3, y = Inf, label = "SFFA ruling",
             color = "grey45", hjust = 1, vjust = 1.5, size = 3.4) +
    scale_x_continuous(breaks = seq(2008, 2024, 4),
                       limits = c(2005.7, 2025.5), expand = c(0, 0)) +
    scale_y_continuous(labels = y_format) +
    coord_cartesian(clip = "off") +
    labs(title = title, x = NULL, y = NULL) +
    theme_minimal(base_size = 12) +
    theme(
      panel.grid          = element_blank(),
      axis.line.x         = element_line(color = "grey55", linewidth = 0.35),
      axis.ticks.x        = element_line(color = "grey55", linewidth = 0.35),
      axis.ticks.y        = element_line(color = "grey55", linewidth = 0.35),
      axis.ticks.length   = unit(3, "pt"),
      axis.text           = element_text(color = "grey30", size = 12),
      plot.title          = element_text(face = "bold", size = 19,
                                         margin = margin(b = 14)),
      plot.title.position = "plot",
      plot.margin         = margin(6, 52, 6, 6)
    )
}
```

```{r}
#| echo: false
#| fig-width: 10
#| fig-height: 5.5
#| label: fig-black
#| fig-cap: "US Black share of the entering class at the forty-seven institutions in the 2026 US News top-25 National Universities and top-25 National Liberal Arts Colleges (service academies excluded), fall 2006 through fall 2025. Grey lines are individual institutions, the orange line is the cross-sectional median, and the dashed vertical line marks the SFFA ruling. The fall 2025 median is computed over the thirty-nine institutions with usable filings. Sources: IPEDS Fall Enrollment and CDS item B2."
spaghetti50(
  pool50 |> filter(!is.na(black)) |> mutate(y = black / us_total),
  title    = "Black Share of First-Year US Students, Top-Ranked Institutions",
  y_format = label_percent(accuracy = 1))
```

The median share occupied a band between 6.0 and 7.7 percent from 2006
through 2019, rose to a maximum of 7.9 percent in fall 2021, and registered
5.9 and 5.8 percent in the two post-ruling cycles. The net movement from
the center of the long-run band is under one percentage point, a fifth or
less of the three-and-a-half to five points that declines of 50 to 70
percent imply on a base near 7 percent. Selection into fall 2025
disclosure does not manufacture this stability. The eight cohort schools
without a usable 2025 filing entered fall 2024 with a median Black share
of 6.8 percent, against 5.4 percent for the thirty-nine that later
filed, so the missing schools, if anything, depress the observed 2025
median. The full-panel figure will likely be higher.

The 2020-2022 increase identifies the presence of discretionary
control. Between the fall 2020 and fall 2022 entering
classes, the median Black share rose more than a full percentage point,
with relative increases above 20 percent at many institutions, over an
interval in which the academic quality and racial composition of the national applicant pool were essentially static. (The College Board's
annual Total Group reports for these cohorts show group means and shares
nearly unchanged; College Board 2020, 2021, 2022.) The proximate cause
was not a change in applicant characteristics but an institutional
response to the events of 2020. The episode establishes that these
institutions possessed, and exercised, the capacity to shift Black
enrollment by large amounts within one to two cycles. The relevant
question for the post-ruling period is whether that capacity was
surrendered or redirected.

@fig-ratio examines the ratio of Asian to Black first-year US students, a
statistic chosen for two properties. It requires no denominator, being a
quotient of two counts from adjacent lines of the same filing, and it is
therefore immune to the international-share and nondisclosure issues
discussed above so long as nondisclosure is not strongly race-selective.
The litigation context also makes it the natural summary measure, since
the plaintiffs were Asian American and the preferences at issue operated
most heavily between the Asian and Black margins.

```{r}
#| echo: false
#| fig-width: 10
#| fig-height: 5.5
#| warning: false
#| label: fig-ratio
#| fig-cap: "Ratio of Asian to Black first-year US students at the same forty-seven institutions. Institution-years with fewer than five Black first-years are excluded. Grey lines are individual institutions, the orange line is the cross-sectional median, and the dashed vertical line marks the SFFA ruling. The four labeled institutions, drawn in darker grey, are Berkeley and UCLA, whose admissions California's Proposition 209 (adopted 1996) has bound since the fall 1998 entering class, and two STEM-intensive schools, Caltech and Carnegie Mellon. Sources: IPEDS Fall Enrollment and CDS item B2."
ratio_df <- pool50 |> filter(!is.na(black), !is.na(asian), black >= 5) |>
  mutate(y = asian / black)
# Right-margin labels for the four schools named in the text. Label y
# positions are each school's final observed ratio (Berkeley 20.8 and
# CMU 13.7 in 2025, UCLA 9.9 in 2025, Caltech 7.5 in 2024). Revisit
# positions if clean.csv is updated.
ratio_labels <- tibble::tibble(
  school = c("berkeley", "cmu", "ucla", "caltech"),
  label  = c("Berkeley", "Carnegie Mellon", "UCLA", "Caltech"),
  y_lab  = c(20.8, 13.7, 9.9, 7.5))
spaghetti50(
  ratio_df,
  title    = "Asian / Black Ratio of First-Year US Students, Top-Ranked Institutions",
  y_format = label_number(accuracy = 1)) +
  geom_line(data = ratio_df |> filter(school %in% ratio_labels$school),
            color = "grey55", linewidth = 0.45, na.rm = TRUE) +
  annotate("text", x = 2025.45, y = ratio_labels$y_lab,
           label = ratio_labels$label, color = "grey35",
           hjust = 0, size = 3.5) +
  theme(plot.margin = margin(6, 112, 6, 6))
```

The cohort median drifted from approximately 1.9 in 2006 to 2.6 by 2023
and averages 3.1 across the two post-ruling cycles. Berkeley and
UCLA, the California institutions bound by Proposition
209, together with the STEM-intensive schools whose applicant pools skew
heavily Asian, occupy ratios between 6 and 13 at the median and up to 25.8
in Berkeley's case (2,396 Asian against 93 Black first-year US students in
2017). California's applicant pool contains proportionally more Asian and
fewer Black students than the national pool serving the private
institutions, and the University of California has admitted test-blind
since 2021, so these values mark the general region rather than the
exact level that race-blind selection would imply elsewhere. The
post-ruling movement of the cohort median, roughly half a point,
covers a small fraction of the interval separating it from that region.

Individual trajectories within the cohort followed a common sequence. The
first post-ruling cycle produced large single-school deviations, with
Stanford rising to 7.4 against a historical range of 2 to 3, Cornell to
6.8, Amherst to 7.0, Brown to 5.0, and Johns Hopkins to
11.8. In the second cycle most of these reverted, Amherst to 2.0, Brown to
3.3, and Cornell to 6.0, while Stanford continued to 9.7 and
Johns Hopkins remained at 11.2. The modal institution experimented with a
substantially higher ratio for one cycle and then returned to its
historical position.

### Two Institutional Trajectories

The aggregate stability documented above conceals two exceptions, one at
each end of the range of institutional response.

Amherst's forecast came true, for exactly one cycle. The brief's one
institution-specific estimate, drawn from the college's internal
planning, concerned combined Native American, Black, and Hispanic
enrollment. That combined share averaged 25.2 percent of Amherst's
entering class over the pre-period and stood at 23.2 percent in fall
2023, the final cohort admitted under racial preferences. The fall 2024
class registered 13.5 percent, or 0.54 of baseline, with the Black
component alone falling from an 11.8 percent baseline to 2.9 percent.
The college's public posture treated the outcome as vindication. It
announced that 3 percent of the Class of 2028 identified as Black under
federal reporting rules, against 11 percent the year before, with
domestic students of color falling from 47 to 38 percent of the class
(Amherst Student 2024a). National coverage presented the numbers as the
realization of the college's own brief. President Michael Elliott called
the outcome disappointing and warned that the change could "diminish the
sense of belonging" of enrolled students (Amherst Student 2024a). The
campus discussion also reproduced, in miniature, the measurement problem
documented for Harvard above: under self-identification the same class was 9 percent
Black, and under the federal single-race rules it was 3 percent, a gap the
student newspaper examined at length (Amherst Student 2024b). The
next cycle reversed the outcome. The college expanded digital and
in-person recruiting, joined the STARS rural-recruitment network, admitted
a record share of first-generation students, and released its class data a
month later than usual (Amherst College 2025; Amherst Student 2025). The
combined underrepresented share returned to 21.2 percent, the Black share
to 5.8 percent, and the Asian-to-Black ratio traced the same arc, rising
to 7.0 in the first post-ruling cycle and returning to 2.0 in the second.
The institution that treated the ruling as binding matched its own
forecast for exactly one cycle and then went back toward the typical racial composition.

Johns Hopkins, by contrast, has presented itself as complying, and its
filings support the presentation. Its announced Class of 2028
figures showed underrepresented enrollment falling from 37.0 to 17.6
percent of the class, with the Black share dropping from 13.8 to 5.7
percent in the university's self-identified categories (JHU News-Letter
2024). President Ronald Daniels responded that the university would pursue
diversity "while fully complying with the Supreme Court's restriction on
the consideration of race in admissions" (Daniels 2024). The second cycle
extended the movement instead of reversing it. Local reporting found
Asian students rising from 26 percent of the entering class in 2023 to 41
percent in 2024 and 45 percent in 2025, a trajectory without parallel
among peer institutions, and the university's comment was that it
"complies with federal law and does not consider race or ethnicity in
admissions" (Baltimore Banner 2026). In my metrics, Hopkins's Black share of US
first-years fell from a 6.0 percent pre-period mean to 4.4 percent, and
its Asian-to-Black ratio registered 11.8 and 11.2 in the two post-ruling
cycles, leaving it, with Stanford, one of only two institutions still
outside its historical range in the second year. Where a school actually
stopped considering race, the movements were large and immediate, and at
Hopkins they persisted, above all in the Asian share. Nothing similar
appears at the forty-five institutions other than Hopkins and Stanford.

### The Test-Taking Population

In the absence of racial preferences, the racial composition of an
elite entering class should, all else equal, resemble the racial
composition of the students who meet elite academic standards. National
testing data describe that population. The College
Board's summary of the 2,004,965 SAT takers in the high school graduating
class of 2025 reports mean total scores of 1,229 among the 217,459 Asian
test-takers and 904 among the 250,887 Black test-takers, a
difference exceeding 300 points against group standard deviations near
200 (196 and 212 for the Black and Asian distributions,
estimated from the report's score-band tables) (College Board 2025). Distributions separated by that
distance thin at very different rates in the upper tail. In the
1400-to-1600 band, approximately the range from which the institutions
studied here draw most of their admits, the report counts 56,539 Asian,
52,079 White, 10,752 Hispanic, and 2,509 Black test-takers (College
Board 2025, score-distribution tables by race and ethnicity), an
Asian-to-Black ratio of 22.5 between two groups of comparable overall
size. No party to the litigation proposed admission by test score alone,
and holistic review weighs more than scores, so the top-band ratio is a rough guide rather than a point prediction. But any selection process
that weights academic credentials heavily and applies no racial offsets
would land far closer to that guide than to the enrolled ratio of 2 to
3, which is little changed after *SFFA* and is not the output of such a
process.

## Conclusion

<!-- The numbers in the conclusion are hard-coded for readability; every
     one restates a figure computed in the body (tbl-fifteen, tbl-amici,
     fig-ratio and surrounding prose). Keep them in sync, and revisit if
     clean.csv is updated. -->
Two admissions cycles after *SFFA v. Harvard*, the enrollment records of
the institutions most affected by the decision show the following:

- Among the 33 colleges that jointly
  forecast declines of 50 to 70 percent, mean Black share is marginally
  higher than in the corresponding benchmark period, and thirty-one of them
  never approached the mildest version of their own forecast.

- At the 15 research universities
  that assured the Court that race-neutral admissions would undercut their
  diversity, the mean Black share of US first-years moved from 7.4 to 7.2
  percent, and six signatories sit above their baselines.

- Across the 47 top-ranked institutions, the ratio of
  Asian to Black first-years, the summary statistic most relevant to the
  underlying litigation and the one least vulnerable to measurement
  objections, moved from roughly 2.6 to roughly 3.1 against race-blind
  benchmarks two to eight times higher.

- Harvard's Black enrollment share, expressed in any internally consistent
  metric, equals its level in the years when it used the admissions system
  the Supreme Court found unlawful.

The parsimonious reading of this
evidence is that elite institutions have not meaningfully changed the
admissions policies the ruling was intended to end. That characterization is an
interpretation of the pattern, not a finding the data compel. The data
cannot identify the mechanism. Whether present practice reflects
consideration of race as such, which the Court prohibited, or of
experiential factors correlated with race, which the Court permitted, is
not answerable from aggregate filings, and I do not claim otherwise.

What
the analysis does establish is that the quantitative expertise deployed in
the litigation has functioned, after the fact, to obscure the question. The most cited prediction in the case was stated
in units that no public statistic reports, by an expert whose analysis
never faced peer review but was certified as sound to the Supreme Court
by sixteen economists including three Nobel laureates, and any
comparison of it against filings
computed under incompatible definitions generates an appearance of
compliance that a factor-of-two conversion dissolves.

Economists who
supply litigation estimates in nonpublic units might reasonably be
expected to publish the mapping to public categories at the same time.
Absent such mappings, predictions of this kind are unfalsifiable in
practice, whatever their merits in principle, and the institutions they
concern are the primary beneficiaries of the ambiguity.

## Data and Code {.unnumbered}

Data and code sufficient to reproduce every number, table, and figure in
this paper, including the extraction pipeline that assembles the Common
Data Set and IPEDS filings for every school in the analysis, are
available at
[https://github.com/davidkane9/after-sffa](https://github.com/davidkane9/after-sffa).

## References {.unnumbered}

Amherst Brief. 2022. Brief of Amici Curiae Amherst, Barnard, Bates,
Bowdoin, Bryn Mawr, Carleton, Colby, Connecticut, Davidson, Franklin &
Marshall, Hamilton, Hampshire, Haverford, Macalester, Middlebury, Mount
Holyoke, Oberlin, Pomona, Reed, Sarah Lawrence, Smith, St. Olaf,
Swarthmore, Trinity, Union, Vassar, Wellesley, and Williams Colleges, and
Bucknell, Clark, Tufts, Washington & Lee, and Wesleyan Universities,
*Students for Fair Admissions v. Harvard*, Nos. 20-1199 and 21-707 (U.S.
Supreme Court, August 1).
[Link](https://www.supremecourt.gov/DocketPDF/20/20-1199/232512/20220801174102841_20-1199%20and%2021-707_Brief%20of%20Amici%20Curiae%20Amherst%20et%20al%20Colleges%20and%20Bucknell%20et%20al%20Universities.pdf)

Amherst College. 2025. Amherst Admits Record Percentage of
First-Generation Students in Class of 2029. Press release, March.
[Link](https://www.amherst.edu/news/news_releases/2025/march/amherst-admits-record-percentage-of-first-generation-students-in-class-of-2029)

Amherst Student. 2024a. Students and Administrators Reckon with a Less
Racially Diverse First-Year Class.
[Link](https://amherststudent.com/article/community-reckons-with-less-diverse-first-year-class/)

Amherst Student. 2024b. Differing Definitions: Making Sense of the
Demographic Data.
[Link](https://amherststudent.com/article/differing-definitions-making-sense-of-the-demographic-data/)

Amherst Student. 2025. College Releases First-Year Class Data After
One-Month Delay.
[Link](https://amherststudent.com/article/college-releases-first-year-class-data-after-one-month-delay/)

Arcidiacono, Peter, Josh Kinsler, and Tyler Ransom. 2022a. Asian American
Discrimination in Harvard Admissions. *European Economic Review* 144:
104079. [Link](https://doi.org/10.1016/j.euroecorev.2022.104079)

Arcidiacono, Peter, Josh Kinsler, and Tyler Ransom. 2022b. Legacy and
Athlete Preferences at Harvard. *Journal of Labor Economics* 40(1):
133-156. [Link](https://doi.org/10.1086/713744)

Arcidiacono, Peter, Josh Kinsler, and Tyler Ransom. 2023. What the
Students for Fair Admissions Cases Reveal About Racial Preferences.
*Journal of Political Economy Microeconomics* 1(4): 615-668.
[Link](https://doi.org/10.1086/725336)

Arcidiacono, Peter, and Michael Lovenheim. 2016. Affirmative Action and
the Quality-Fit Trade-off. *Journal of Economic Literature* 54(1): 3-51.
[Link](https://doi.org/10.1257/jel.54.1.3)

Baltimore Banner. 2026. Asian Enrollment at Johns Hopkins Is
Skyrocketing. No One Can Say Why. April 14 (by Ellie Wolfe).
[Link](https://www.thebanner.com/education/higher-education/johns-hopkins-university-asian-enrollment-K7SYLRTSLFB35JBTXDH46BVWMA/)

Bleemer, Zachary. 2022. Affirmative Action, Mismatch, and Economic
Mobility After California's Proposition 209. *Quarterly Journal of
Economics* 137(1): 115-160.
[Link](https://doi.org/10.1093/qje/qjab027)

Bloem, Michael D., Ashley Edwards, J. Parker Goyer, Jessica Howell,
Xiaowen Hu, Michael Hurwitz, Samuel J. Imlay, Jennifer Ma, and Matea
Pender. 2026. College Enrollment Patterns After *SFFA v. Harvard*.
EdWorkingPaper 26-1392, Annenberg Institute, Brown University (January).
[Link](https://doi.org/10.26300/6a7w-bq06)

Bowen, William G., and Derek Bok. 2018. *The Shape of the River:
Long-Term Consequences of Considering Race in College and University
Admissions*. 20th anniversary ed. Princeton: Princeton University Press.
First published 1998.

Card, David. 2017. Report of David Card, Ph.D., *Students for Fair
Admissions v. Harvard*, No. 1:14-cv-14176 (D. Mass., December 15).

Chetty, Raj, David J. Deming, and John N. Friedman. 2026. Diversifying
Society's Leaders? The Determinants and Causal Effects of Admission to
Highly Selective Private Colleges. *Quarterly Journal of Economics*
141(1): 51-145.
[Link](https://academic.oup.com/qje/article-abstract/141/1/51/8306880)

College Board. 2020. *2020 Total Group SAT Suite of Assessments Annual
Report*. New York: College Board.
[Link](https://reports.collegeboard.org/media/pdf/2020-total-group-sat-suite-assessments-annual-report.pdf)

College Board. 2021. *2021 Total Group SAT Suite of Assessments Annual
Report*. New York: College Board.
[Link](https://reports.collegeboard.org/media/2022-04/2021-total-group-sat-suite-of-assessments-annual-report%20%281%29.pdf)

College Board. 2022. *2022 Total Group SAT Suite of Assessments Annual
Report*. New York: College Board.
[Link](https://reports.collegeboard.org/media/pdf/2022-total-group-sat-suite-of-assessments-annual-report.pdf)

College Board. 2025. *2025 Total Group SAT Suite of Assessments Annual
Report*. New York: College Board.
[Link](https://reports.collegeboard.org/media/pdf/2025-total-group-sat-suite-of-assessments-annual-report%20ADA-v0.2_0.pdf)

Daniels, Ronald J. 2024. Johns Hopkins' Continued Commitment to
Diversity. Message from the Office of the President, Johns Hopkins
University, September 19.
[Link](https://president.jhu.edu/messages/2024/09/19/johns-hopkins-continued-commitment-to-diversity/)

Economics Professors Brief. 2022. Brief of Professors of Economics as
Amici Curiae in Support of Respondent, *Students for Fair Admissions v.
Harvard*, No. 20-1199 (U.S. Supreme Court, August 1).
[Link](https://www.supremecourt.gov/DocketPDF/20/20-1199/232347/20220801133551596_Economics%20Profs.%20amicus%20brief%20-%20SFFA%20v.%20Harvard%20-%20No.%2020-1199.pdf)

Economists Brief. 2022. Brief of Economists as Amici Curiae in Support of
Petitioner, *Students for Fair Admissions v. Harvard*, No. 20-1199 (U.S.
Supreme Court, May 9).
[Link](https://www.supremecourt.gov/docket/docketfiles/html/public/20-1199.html)

Espenshade, Thomas J., and Chang Y. Chung. 2005. The Opportunity Cost of
Admission Preferences at Elite Universities. *Social Science Quarterly*
86(2): 293-305.
[Link](https://doi.org/10.1111/j.0038-4941.2005.00303.x)

Grodsky, Eric, and Michal Kurlaender. 2006. The Demography of Higher
Education in the Wake of Affirmative Action. Working paper, University of
California, Davis (October). [Link](https://tinyurl.com/2p9eda8c)

Harvard Crimson. 2025. Black and Hispanic Enrollment Drops, Asian
Enrollment Rises for Harvard's Class of 2029. October 23.
[Link](https://www.thecrimson.com/article/2025/10/23/admissions-data-class-2029/)

Harvard Gazette. 2024. Harvard Releases Race Data for Class of 2028.
September 11.
[Link](https://news.harvard.edu/gazette/story/2024/09/harvard-releases-race-data-for-class-of-2028/)

Hinrichs, Peter. 2012. The Effects of Affirmative Action Bans on College
Enrollment, Educational Attainment, and the Demographic Composition of
Universities. *Review of Economics and Statistics* 94(3): 712-722.
[Link](https://doi.org/10.1162/REST_a_00170)

JHU News-Letter. 2024. University Announces Drop in Percentage of
Students from Underrepresented Groups in the Class of 2028. *Johns
Hopkins News-Letter*, September 21.
[Link](https://www.jhunewsletter.com/article/2024/09/university-announces-drop-in-percentage-of-students-from-underrepresented-groups-in-the-class-of-2028)

Universities Brief. 2022. Brief of Brown University, California Institute
of Technology, Carnegie Mellon University, Columbia University, Cornell
University, Dartmouth College, Duke University, Emory University, Johns
Hopkins University, Princeton University, University of Chicago,
University of Pennsylvania, Vanderbilt University, Washington University
in St. Louis, and Yale University as Amici Curiae, *Students for Fair
Admissions v. Harvard*, Nos. 20-1199 and 21-707 (U.S. Supreme Court,
August 1).
[Link](https://www.supremecourt.gov/DocketPDF/20/20-1199/232422/20220801150520881_20-1199%20%2021-707%20bsac%20Universities.pdf)

## Cases Cited {.unnumbered}

*Grutter v. Bollinger*, 539 U.S. 306 (2003).

*Students for Fair Admissions v. President and Fellows of Harvard
College*, 397 F. Supp. 3d 126 (D. Mass. 2019).

*Students for Fair Admissions v. President and Fellows of Harvard
College*, 600 U.S. 181 (2023).

## About the Author {.unnumbered}

David Kane (www.davidkane.info) is a visiting assistant professor at New College of Florida.
He is the former Preceptor in Statistical Methods and Mathematics in the
Department of Government at Harvard University. He served as an officer
in the United States Marine Corps from 1988 to 1991. Prior to returning
to academia in 2018, he worked in quantitative finance for two decades,
mostly as an equity portfolio manager at firms including Numeric, Geode,
Millennium, Point72 (formerly SAC), Acadian, and Hutchin Hill. His email
is dave.kane@gmail.com.

